\chapter{Language Design}
\label{cha:Design}
%Metatext for design chapter
In this chapter the design of the language called \gls{gamble} will be presented.\todo{Hvad står Gamble for?}
First with a brief description of our design philosophy for \gls{gamble} will be introduced to clarify what this language is attempting to achieve and why.
The design philosophy will address some of the criteria found in \myref{sec:OurCriterias}.
Furthermore the features and design of \gls{gamble} will be documented later in this chapter.

\input{chapters/sections/Philosophy.tex}

In the following sections different parts of \gls{gamble} will be investigated.

\section{Core syntax choices}
The source code must be encoded in UTF-8, but only the ASCII alphabet and numbers are allowed for identifiers and values, this is to simplify the parsing. 
The language is case sensitive, because the most used languages seen on TIOBEs index\citep{TIOBE} are also case sensitive.
When a language is case sensitive it is possible for the programmer to name different parts of the code like he pleases, resulting in higher readability.
Some languages, such as Python, uses whitespace and indentation to indicate scope, in our language scope is indicated by curly brackets like many existing languages\texttt{\{\}}.  

\subsection*{Comments}
It is useful to be able to write comments in code to annotate the meaning of some code.
If a language were not to have the possibility to write comments, then the source code could be hard to understand.
Source code without comments might be difficult to return to after some time away from it, since the programmer would have to reread the entire code, instead of a few good comments.\citep{Commenting}
Therefore both single line and multi line comments can be used in \gls{gamble}. 
``//'' is used for single line comments, meaning that everything after the ``//'' until the next newline is ignored by the compiler, also known as C++ style comments. 
``/* */'' is used for multi line comments, meaning that everything between ``/*'' and ``*/'' is ignored by the compiler, also known as C style comments. 

\subsection*{End of statement terminator}
When choosing an end of state terminator there are two main choices.
The newline character as a terminator or another symbol, often semicolon (;).
The use of newline is often used in simple languages e.g. MATLAB or other mathematical oriented languages. 
We deem the main value of using the newline are simplicity and a decrease in errors due to forgotten semicolons.
Semicolon is as mentioned the other main choice, this choice has been used through many programming languages throughout the history. 
The main advantage of a end of statement terminator  like semicolon is that it allows freeform code. 
This freeform allows to have multiple statements in one line or allows long statements to span across multiple lines.
We deem this to enhance the readability of the code.
Furthermore the semicolon terminator allows for a simpler syntax and compiler design.
As mentioned before forgotten semicolons can result in compiler errors, we deem this as the biggest disadvantages of this end of statement terminator.

Based on this analysis we have chosen to use the semicolon as end of statement terminator, for easier parsing.

\subsection*{Scope}\label{subsec:Scope}
If a programming language contain no implementation of scoping rules, then one must assume that all the source code in a given program would be in the same scope.
Such a single scope would mean that any variable can be used form any place in the code. 
Mostly, as briefly described in \myref{sec:state_of_the_art}, a programming language implements some kind of scoping rules.
There are several reasons for this, the main reason being it gives the programmer more control of the code and reduces the chance of changing variables by accident.
There exist two main methods of scoping, static and dynamic scoping.
As seen in \myref{sec:state_of_the_art} most of the languages examined in this section used a form of static scope. 
\gls{gamble} will use this form of scoping because we find it to be the easiest to work with and less prone to errors than dynamic scoping.
The scope rules of \gls{gamble} are as follows:

A variable is in scope from its declaration until the end of the block it is declared in.
An inner block such as a control flow structure, see \myref{subsec:control-flow} for more, has access to the variables from the outer scope. 
It is also not allowed to redeclare a variable which is already declared in the current scope or any enclosing scopes. 
Additionally any function is in scope of any other function including itself, this also allows recursion. 
- This is static scoping without information hiding. 

\subsection*{Structure}\label{subsec:Struc}
The structure of the source code to a \gls{gamble} program is shown in \myref{lst:Structure}.
For a high level of readability we have decided  to split the source code up in segments.
A single segment contains only a small part of source code.
Including of other files are, like in many other languages, written in the top segment, this is to make it easy to get an overview of any and all included libraries. 
All function declarations are written in the next segment followed by the main statements which is the last segment.
This collects the main statements, which makes it easier for a programmer to get an overview over the order of execution, compared to if the main statements were allowed between the function declarations.
This sectioning also makes it easier to write a context free grammar for the language.

\begin{lstlisting}[caption={Source code file layout in \gls{gamble}},frame=tlrb,label={lst:Structure}, numbers=none]
Libary inclusions

Function declarations

Statements
\end{lstlisting}

\section{Types and Variables}\label{sec:Types}
%http://www.informit.com/articles/article.aspx?p=2103809&seqNum=3
\gls{gamble} uses three primitive data types; integers (int), floating points (float) and boolean values (bool), as well as two composite data types; matrices and vector. 
In \gls{gamble} any variable must be initialised when it is declared. 
This avoids any null reference which can cause runtime errors in many languages such as C, C\# and Java.
In \gls{gamble} all variables are statically stored, there are several reasons for this and therefore the language does not implement neither explicit or dynamic memory handling.
The first reason for this is that the purpose of the language is to perform arithmetic operations on the \acrshort{gpu}, so by excluding this the language's simplicity increases, and therefore its read- and writability increases.
A second reason being that the size of every variable in \gls{gamble} is unchangeable.
Any vector and matrix will be allocated on the heap by the compiler.
Manual memory handling could increase the number of possible programs in the language and is therefore seen as a viable extension to future versions of \gls{gamble}.


\subsection*{Integers and floating points}
An integer is a whole number either positive or negative, represented by the two's compliment method. 
Meaning that it ranges from $-(2^{n-1}) $ to $2^{n-1} - 1 $, where n is the number of bits used.
A floating point number is an approximation to a real number. 
It is separated in 3 parts, the sign bit, the mantissa and the exponent. 
Floats in \gls{gamble} are in base 2, the value of a float is calculated by $ {(-1)}^{sign bit} \cdot mantissa \cdot 2^{exponent} $. 
Integers and floating point numbers default to 32 bits with the keywords \texttt{int} and \texttt{float}. 
It is possible to declare 16 and 64 bit integers and floating point numbers by postfixing their size to their respective keywords, e.g. \texttt{float64} or \texttt{int16}. 
It is useful to be able to specify the size of an integer or a floating point number, as data transfer can cause a lot of overhead particularly when transferring data to and from the \acrshort{gpu}. 

\subsection*{Boolean}
A boolean value is either true or false. 
These are used for control structures in \gls{gamble} which takes a boolean value as a parameter. 
But can also be used for any means which the programmer finds useful. 

\subsection*{Vectors and Matrices}
A vector in \gls{gamble} is a sequence of entries which are each the same primitive data type. 
The dimension of a vector is immutable, meaning that its dimension cannot be changed after its declaration. 
A matrix is multiple row vectors inside a column vector, each row vector is the same dimension. 
As for other variables both vectors and matrices must be initialised when they are declared. 
This is done with the syntax shown in \myref{lst:matrix}.
As seen there are two ways of declaring a matrix and a vector, one can either provide the values for each entry, or one can provide the dimensions required of a matrix or vector in which case all values are initialised as zero.
The syntax for the latter does not adhere to the concept of orthogonality mentioned in \myref{cha:language_criteria} although the use of this syntax should be held to a minimum, as it primarily exists such that functions in the standard library for \gls{gamble} can be created.
What these functions entail is described in \myref{sec:funcs}.
Here it is also seen why the use of ``;'' is a smart choice for ending statements.\todo{what does this mean?}

\begin{lstlisting}[caption={Syntax for creating a matrix or vector},label={lst:matrix},numbers=none]
//Makes a 2x2 matrix by providing entry values
matrix<int> m = [1, 0; 
                 0, 1];
//Makes a 100x100 matrix with all values set to 0
matrix<int>[100, 100] mm;

//Making a vector by providing values
vector<int> v = [1, 0];
//Makes a vector with 100 entries all set to 0
vecotr<int>[100] vv;
\end{lstlisting}

\subsection*{Operators}
Available operators for simple data types in the language can be found on \myref{tbl:operators} with a short description.  

%It is not possible to use operators with matrices or vectors, functions which support different operations on these types will be included in a library.
\input{figures/Operators.tex}
Some operators will work work with vectors and matrices.
Since the elements in vectors and matrices are one of the simple types of \gls{gamble}, the operands on both sides of an infix operators must adhere to the rules of type compatibility.
The list of possible operands on these complex data types can be seen on \myref{tbl:matOps}.
\input{figures/matrixOperators.tex}
%\subsection*{Memory handling}
%Locally declared and small and simple variables will be written to the stack, while large quantities of data must be written to the heap. 
%To manage the heap in a programming language one must decide how the management must be implemented.
%At the topmost level there exist to different approaches, the first being explicit memory handling as seen in C with function calls like \texttt{alloc()} and \texttt{free()}. 
%The second approach is dynamic memory handling, often called garbage collection.
%While  manual memory handling is a very low level concept, it has some advantages over dynamic memory handling, for instance manual handling gives the users full control over a programs resources and while this can be hard o get right, can both increase the simplicity of the compiler and increase a programs runtime execution speed because garbage collection can be a complex runtime process.\citep{Sebesta, gribble}

\section{Functions}\label{sec:funcs}
Functions are blocks of code which can be called to do a specific computation, often but not always, a function takes some variables as inputs.

This section will describe \gls{gamble}'s use of functions. 
Like in many other languages it is possible to declare your own functions in \gls{gamble}.
This is useful for organising code, and reusing parts of the source code.
Functions in \gls{gamble} can take input, but are not required to, the inputs can be any of the types the language support.
Because of high implementation cost in the compiler, it is not possible to use a function as an input for a function.

\gls{gamble} also have certain functions built into the language for varying reasons.
The most unique function in question here will be the \texttt{print} function.
This function is unique in the sense that it enables the programmer to use strings as a parameter, a data type otherwise not implemented in the language.
This is enabled such that the programmer can use this feature to write out messages to the console. 
This is both useful for output and debugging the source code.

Other functions provided by \gls{gamble} are created out of convenience.
In object-oriented languages subscripting is common and allows for a wide range of useful features, e.g. \texttt{.GetLength} on arrays in C\#.
Similar to this \gls{gamble} provides \texttt{Rows} and \texttt{Cols} functions for matrices and vectors which returns their amount of rows or columns.
If called on a vector one can use either rows or columns, it will give the same result.
This can be used in loops to make them iterate over matrices of any size.

Another such function is \texttt{MakeIntMatrix}.
This creates a matrix with int values a similar function exists for float.
These takes three arguments as input, the amount of rows, amount of columns, and what value to place at every entry.
This function builds upon the syntax for initialising a matrix by providing dimensions rather than input, it merely alters the initialised value depending on the argument provided.
Other functions are created out of necessity, functions for arithmetics such as multiply and addition also exist, as operators between matrices are not a valid expression.

\subsection*{Function identifiers}
When declaring a \gls{gamble} function it is required to write the body of the function immediately after.
This is to avoid lookahead which function prototypes in C requires.
A function identifier is exactly like C, with the \texttt{return type} then the \texttt{functionname}, \texttt{(formal parameters)} and finally the \texttt{\{body\}}. % chktex 36
The function body can contain a number of statements and various control flows.
Furthermore it can contain calls to other functions or calls to the function itself, and hereby recursion is a possibility in \gls{gamble}
An example of a function identifier can be seen on \myref{lst:functionID}.
Other syntaxes could have been used, but not only is this the way C does it, it is also widely used in many other languages i.e. C\# and Java.

\begin{lstlisting}[caption={Function Identifier},label={lst:functionID},numbers=none]                                                        
int add(int a, int b)
{
    return a + b;
}
\end{lstlisting}

\subsection*{Function calls}
After a function has been declared it can be called. 
A function call contains the identifier and parameters for the function. 
The syntax in \gls{gamble} is the same as in many other programming language and shown in \myref{lst:functionCall}. 

\begin{lstlisting}[caption={A function call in \gls{gamble}},label={lst:functionCall},numbers=none]
add(4, 3);
\end{lstlisting}


\subsection*{Return value}
Functions have a return value which can be seen on \myref{lst:functionID}.
The return values are all the types in \gls{gamble} mentioned in \myref{sec:Types} and void. 
A void function will not return anything, but instead can be seen as a procedure.
This choice was made because \gls{gamble} should be able to return all the different types in the language from a function, but also be able to perform these procedures which a void function may do.
The value or type to be returned is preceded by the keyword \texttt{return}, just like it is in C and other C-like languages.
An example can be seen on \myref{lst:returnFunction}.

\begin{lstlisting}[caption={Return Function},label={lst:returnFunction}]
int a = 0;

a = add(4, 3);
\end{lstlisting}


%\subsection*{Print function}\todo{rename maybe}\todo{skal flyttes op til de andre funktioner?}
%In \gls{gamble} we have one function that is ready-made. It is a print function.
%The function is special because it can both take variables, numerical values and strings as input parameter.
%Since string is not a valid data type in \gls{gamble} this function is special. 
%With the print function it is possible to give the user a visual representation of the computations result, in the console.
 
%STD LIB EVT?!
\input{chapters/sections/subsections/controlflow2.tex}
\input{chapters/sections/subsections/reservedwords.tex}