\chapter{Semantics}\todo{Der skal eventuelt skrives noget omkring hvad det semantic der er sat ind repræsentere og hvordan det foregår? - Marc - Enig, der bør nok være lidt teori egentlig, der viser en dummy rule eller noget - Søren}
In this section some of the semantics for \gls{gamble} will be shown and explained.
Operational semantics are provided to formalise information about the mathematical operations behind the language.
There is two ways of showing operational semantics natural semantics and structural operational semantics.
Natural semantics often called big-step semantics and structural operational semantics often referred to as small-step semantics both shows the mathematical rules of a language. 
In this chapter big-step semantics are shown which, as implied by it's name, describe execution of operations through overall results.

The names of some production rules of \gls{gamble} are changed to simplify some of the rules as follows:

\begin{align*}
	\textbf{a} \in  &Arithmetic\quad expression\\
	\textbf{v} \in  &Variable\\
	\textbf{m} \in  &Matrix\\
	\textbf{vec} \in  &Vector\\
	\textbf{type} \in &Int | Float | Bool\\
	\textbf{S}	\in &Statement\\
	\textbf{mtype} \in &Int | Float\\
	\textbf{Fd}		\in &Function Declaration
\end{align*}

\textbf{v} is meant to store the result, this can be both a value, a vector or a matrix.
The big-step semantic operations for a matrix, also shows the semantics for a vector, which evaluates to a matrix with one column.
On the other hand, big-step semantic operations for a vector are not applicable to a matrix, since it requires the dimensions to fit those of a matrix with one column.

The variable environment(${ e }_{ v }$) is a function that for each variable defines which storage location it is bound to. A variable environment corresponds to a table containing symbols which are in that environment.
The store ($st$) is a function that for each storage location in the memory tells us which value is stored in the location. A store corresponds to a complete description of the contents of the memory.\citep{EnvSt_Semantics}

A few of the big-step semantics for \gls{gamble} is included in this chapter, while the rest of the semantics for \gls{gamble} can be found in \myref{app:semantics}.
Some of the operational semantics which are covered in this chapter are variable declaration, assignment, matrix declaration, matrix multiplication and scalar product.

\subsection*{Declaring Variables}
Variables declarations are also statements according to \gls{gamble}'s grammar.
Declarations have an effect on the variable environment which is why it has it's own section.
A variable declaration modifies the store and the variable environment, the new location in the store will be initialised to contain the values of the new variables, and the new variable will exist in the variable environment.

\todo{der mangler da vidst et komma heri}
[${VariableDeclaration}_{BSS}$]
\begin{equation}
	\frac { <S,{ e }_{ v }^{ `` },st[\iota \mapsto v]>{ \rightarrow  }_{ S }<{ e }_{ v }^{ ` },{ st }^{ ` }> }{ <type\quad x=a;S,{ e }_{ v }st>{ \rightarrow  }_{ S }<{ e }_{ v }^{ ` },{ st }^{ ` }> } 
\end{equation}
\begin{align*}
	where\quad &{ e }_{ v },st\vdash a{ \rightarrow  }_{ A }v \\
	&\iota {= e }_{ v }(next) \\
	&{ e }_{ v }^{ `` }={ e }_{ v }[x\mapsto \iota ][next\mapsto new \iota ]
\end{align*}

\section*{Statements}
The effect of a statement is that the store may change, since a statement may modify the values of variables involved through assignments.
A statement should not modify the variable environment.\todo{why? - Marc - Tilføj: "because it creates no new instances of variables" - Søren}
We define a big-step semantic for statements, except for procedure calls.

[${Assignment}_{BSS}$]
\begin{align*}
	&{ e }_{ f },{ e }_{ v }\vdash <x=a,st>\rightarrow st[\iota \mapsto v]\\
	where\quad &{ e }_{ v },st\vdash a{ \rightarrow  }_{ A }v\quad and\quad  { e }_{ v }(x)=\iota 
\end{align*}

[${While\top}_{BSS}$]
\begin{align*}
	&\frac { { e }_{ f },{ e }_{ v }\vdash <{ S }_{ 1 },st>\rightarrow { st }^{ `` },\quad { e }_{ v }\vdash <while\quad (b)\quad \{S\},{ st }^{ `` }>\rightarrow { st }^{ ` } }{ { e }_{ v }\vdash <while\quad (b)\quad \{S\},st>\rightarrow { st }^{ ` } }\\
	&if\quad { e }_{ v },st\vdash b{ \rightarrow  }_{ B }\top 
\end{align*}


[${While\bot}_{BSS}$]
\begin{align*}
	&{ e }_{ f },{ e }_{ v }\vdash <while\quad (b)\quad \{S\},\quad st>\rightarrow st\\
	&if\quad { e }_{ v },st\vdash b{ \rightarrow  }_{ B }\bot
\end{align*}

\section*{Matrices and Vectors}
A matrix is referred to as seen below, the superscript refers to name of the matrix, while the subscript refers to the entry in the matrix.
\begin{align*}
	M^{n}= \begin{bmatrix} { { m }_{ 1,1 }^{ n } } & { { m }_{ 1,2 }^{ n } } & \dots  & { m }_{ 1,k }^{ n } \\
{ { m }_{ 2,1 }^{ n } }  &  { { m }_{ 2,2 }^{ n } } & \dots & { m }_{ 2,k }^{ n }
\\ \vdots  & \vdots & \ddots  & \vdots \\
 { m }_{ j,1 }^{ n } & { m }_{ j,2 }^{ n } & \dots & { m }_{ j,k }^{ n } \end{bmatrix}
\end{align*}

[${MatrixDeclaration}_{BSS}$]
\begin{equation}
	\frac { { e }_{ f },{ e }_{ v },st\vdash j{ \rightarrow  }_{ A }{ v }_{ 1 },\quad { e }_{ f },{ e }_{ v },st\vdash k{ \rightarrow  }_{ A }{ v }_{ 2 } }{ { M }^{ 1 }\quad =\quad matrix<int|float>[{ v }_{ 1 },{ v }_{ 2 }]\rightarrow { st }^{ ` },{ e }_{ v } } ,{ M }^{ 1 }[{ v }_{ 1 },{ v }_{ 2 }]=\begin{bmatrix} { { 0 }_{ 1,1 } } & 0_{ 1,2 } & \dots  & { 0 }_{ { 1,{ v }_{ 2 } } } \\ 0_{ 2,1 } & 0_{ 2,2 } & \dots  & 0_{ 2,v_{ 2 } } \\ \vdots  & \vdots  & \ddots  & \vdots  \\ { 0 }_{ { v }_{ 1 },1 } & 0_{ v_{ 1 },2 } & \dots  & { 0 }_{ { v }_{ 1 },{ v }_{ 2 } } \end{bmatrix}
\end{equation}




[${MatrixMultiplication}_{BSS}$]

Matrix multiplication takes two matrices of the same type, where the size of $m^{1}$ and $m^{2}$satisfy $m^{1}[j,k]*m^{2}[k,l]$.
The operation returns a matrix of with $j$ columns and $l$ rows based on the rules in equation \ref{equ:matrix_multiplicatial}.


\begin{minipage}{1.0\textwidth}
\begin{equation} \label{equ:matrix_multiplicatial} 
\begin{aligned}
	\frac { { e }_{ v },st\vdash { M }^{ 1 }{ \rightarrow  }_{ A }{ v }_{ 1 }\quad { e }_{ v },st\vdash { M }^{ 2 }{ \rightarrow  }_{ A }{ v }_{ 2 } }{ { e }_{ v },st\vdash { M }^{ 1 }*{ M }^{ 2 }{ \rightarrow  }_{ A }{ v } } ,\begin{matrix} { v }_{ 1 }=matrix<int|float>[j,k] \\ { v }_{ 2 }=matrix<int|float>[k,l] \\
	 v=\begin{bmatrix} { { m }_{ 1,1 }^{ 3 } } & { { m }_{ 1,2 }^{ 3 } } & \dots  & { m }_{ 1,l }^{ 3 }\\
{ { m }_{ 2,1 }^{ 3 } } &   +{ { m }_{ 2,2 }^{ 3 } } & \dots & { { m }_{ 2,l }^{ 3 } }
\\ \vdots  & \vdots & \ddots  & \vdots  \\
 { m }_{ j,1 }^{ 3 }& { { m }_{ j,2 }^{ 3 } } & \dots & { m }_{ j,l }^{ 3 } \end{bmatrix} where
\\ \end{matrix}
\end{aligned}
\end{equation}
\begin{equation*}
\begin{aligned}
{ { m }_{ 1,1 }^{ 3 } }  &= { { m }_{ 1,1 }^{ 1 } } *{ { m }_{ 1,1 }^{ 2 } }+ { { m }_{ 1,2 }^{ 1 } } *{ { m }_{ 2,1 }^{ 2 } }+\dots+ { { m }_{ 1,k }^{ 1 } } *{ { m }_{ k,1 }^{ 2 } }\\
{ { m }_{ 1,2 }^{ 3 } }  &= { { m }_{ 1,1 }^{ 1 } } *{ { m }_{ 1,2 }^{ 2 } }+ { { m }_{ 1,2 }^{ 1 } } *{ { m }_{ 2,2 }^{ 2 } }+\dots+ { { m }_{ 1,k }^{ 1 } } *{ { m }_{ k,2 }^{ 2 } }\\
\vdots\\
{ { m }_{ 1,l }^{ 3 } }  &= { { m }_{ 1,1 }^{ 1 } } *{ { m }_{ 1,l }^{ 2 } }+ { { m }_{ 1,2 }^{ 1 } } *{ { m }_{ 2,k }^{ 2 } }+\dots+ { { m }_{ 1,j }^{ 1 } } *{ { m }_{ j,l }^{ 2 } }\\
{ { m }_{ 2,1 }^{ 3 } }  &= { { m }_{ 2,1 }^{ 1 } } *{ { m }_{ 1,1 }^{ 2 } }+ { { m }_{ 2,2 }^{ 1 } } *{ { m }_{ 2,1 }^{ 2 } }+\dots+ { { m }_{ 2,k }^{ 1 } } *{ { m }_{ k,1 }^{ 2 } }\\
{ { m }_{ 2,2 }^{ 3 } }  &= { { m }_{ 2,1 }^{ 1 } } *{ { m }_{ 1,l }^{ 2 } }+ { { m }_{ 2,2 }^{ 1 } } *{ { m }_{ 2,2 }^{ 2 } }+\dots+ { { m }_{ 2,k }^{ 1 } } *{ { m }_{ k,2 }^{ 2 } }\\
&\vdots\\
{ { m }_{ 2,l }^{ 3 } }  &= { { m }_{ 2,1 }^{ 1 } } *{ { m }_{ 1,l }^{ 2 } }+ { { m }_{ 2,2 }^{ 1 } } *{ { m }_{ 2,l }^{ 2 } }+\dots+ { { m }_{ 2,k }^{ 1 } } *{ { m }_{ k,l }^{ 2 } }\\
&\vdots\\
{ { m }_{ j,1 }^{ 3 } }  &= { { m }_{ j,1 }^{ 1 } } *{ { m }_{ 1,1 }^{ 2 } }+ { { m }_{ j,2 }^{ 1 } } *{ { m }_{ 2,1 }^{ 2 } }+\dots+ { { m }_{ j,k }^{ 1 } } *{ { m }_{ k,1 }^{ 2 } }\\
{ { m }_{ j,2 }^{ 3 } }  &= { { m }_{ j,1 }^{ 1 } } *{ { m }_{ 1,2 }^{ 2 } }+ { { m }_{ j,2 }^{ 1 } } *{ { m }_{ 2,2 }^{ 2 } }+\dots+ { { m }_{ j,k }^{ 1 } } *{ { m }_{ k,2 }^{ 2 } }\\
&\vdots\\
{ { m }_{ j,l }^{ 3 } }  &= { { m }_{ j,1 }^{ 1 } } *{ { m }_{ 1,l }^{ 2 } }+ { { m }_{ j,2 }^{ 1 } } *{ { m }_{ 2,l }^{ 2 } }+\dots+ { { m }_{ j,k }^{ 1 } } *{ { m }_{ k,l }^{ 2 } }
\end{aligned}
\end{equation*}
\end{minipage}




[${ScalarMultiplication}_{BSS}$]
A scalar multiplication is a simple operation on a matrix, where every element in the $j,k$ matrix is multiplayed with an Arithmetic expression $a$. 

\begin{equation}
	\frac { { e }_{ v },st\vdash { M }^{ 1 }{ \rightarrow  }_{ A }{ v }_{ 1 }\quad { e }_{ v },st\vdash { a }_{ 1 }{ \rightarrow  }_{ A }{ v }_{ 2 } }{ { e }_{ v },st\vdash { M }^{ 1 }\ast { a }_{ 1 }{ \rightarrow  }_{ A }{ v } } ,\begin{matrix} { v }_{ 1 }=matrix<int|float>[j,k] \\  \\ v=\begin{bmatrix} { { a }_{ 1 }*m }_{ 1,1 }^{ 1 } & { { a }_{ 1 }*m }_{ 1,2 }^{ 1 } & \dots  & { { a }_{ 1 }*m }_{ 1,k }^{ 1 } \\ { { a }_{ 1 }*m }_{ 2,1 }^{ 1 } & { { a }_{ 1 }*m }_{ 2,2 }^{ 1 } & \dots  & { { a }_{ 1 }*m }_{ 2,k }^{ 1 } \\ \vdots  & \vdots  & \ddots  & \vdots  \\ { { a }_{ 1 }*m }_{ j,1 }^{ 1 } & { { a }_{ 1 }*m }_{ j,2 }^{ 1 } & \dots  & { { a }_{ 1 }*m }_{ j,k }^{ 1 } \end{bmatrix} \end{matrix}
\end{equation}
\todo[inline{navne på reglerne sidder helt tilfældigt, det skal vi ha fikset MP}